---
output:
  word_document: default
  html_document: default
---
```{r,echo=FALSE}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(ggplot2)) install.packages("ggplot2", repos = "http://cran.us.r-project.org")
if(!require(caTools)) install.packages("caTools", repos = "http://cran.us.r-project.org")
if(!require(klaR)) install.packages("klaR", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")
if(!require(dplyr)) install.packages("dplyr", repos = "http://cran.us.r-project.org")
if(!require(broom)) install.packages("broom", repos = "http://cran.us.r-project.org")
if(!require(MASS)) install.packages("MASS", repos = "http://cran.us.r-project.org")
if(!require(corrplot)) install.packages("corrplot", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(ggplot2)
library(caTools)
library(klaR)
library(data.table)
library(dplyr)
library(broom)
library(MASS)
library(corrplot)

theme_set(theme_classic())
```

```{r}
data("PimaIndiansDiabetes2", package = "mlbench")
PimaIndiansDiabetes2 <- na.omit(PimaIndiansDiabetes2)
model <- glm(diabetes ~., data = PimaIndiansDiabetes2,
             family = binomial)
probabilities <- predict(model, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "pos", "neg")
head(predicted.classes)
```

```{r}
mydata <- PimaIndiansDiabetes2 %>%
  dplyr::select_if(is.numeric)
predictors <- colnames(mydata)
```

```{r}
#Dim
dim(mydata)
#Summary
summary(mydata)
#Str
str(mydata)
```

```{r}
#Barplots
par(mfrow=c(2,4))
for(i in 1:8) {
  counts <- table(mydata[,i])
  name <- names(mydata)[i]
  barplot(counts, main=name)
}

```

```{r}
correlations <- cor(mydata[,1:8])
corrplot(correlations, method="circle")
corrplot(correlations, method="pie")
corrplot(correlations, method="square")
corrplot(correlations, method="number")
```

```{r}
#Scatter plot
pairs(mydata)
```

```{r}
mydata <- mydata %>%
  mutate(logit = log(probabilities/(1-probabilities))) %>%
  gather(key = "predictors", value = "predictor.value", -logit)
ggplot(mydata, aes(logit, predictor.value))+
  geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth(method = "loess") +
  theme_bw() +
  facet_wrap(~predictors, scales = "free_y")
```

##
Cook’s distance is used to examine the extreme values in the data. Below is the top 3 largest values.
```{r}
plot(model, which = 4, id.n = 3)
```


##
Further details of the top 3 largest values are shown as follow:

```{r}
model.data <- augment(model) %>%
  mutate(index = 1:n())
model.data %>% top_n(3, .cooksd)
```

##
Although influential values (extreme values) may alter the quality of the logistic regression model, not all outliers are influential observations. To check if the data contains potential influential observations, the standardized residual error can be inspected. The standardized residuals are plotted as below.

```{r}
ggplot(model.data, aes(index, .std.resid)) +
  geom_point(aes(color = diabetes), alpha = .5) +
  theme_bw()

```

##
The filter below is used to identify influential data points with abs (.std.res) > 3. The result shows that there is no influential observations in the data.

```{r}
model.data %>%
  filter(abs(.std.resid) > 3)
car::vif(model)
```

```{r}
pima.data <- na.omit(PimaIndiansDiabetes2)
# Inspect the data
sample_n(pima.data, 3)
# Split the data into training and test set
set.seed(123)
training.samples <- pima.data$diabetes %>%
  createDataPartition(p=0.8, list=FALSE)
train.data <- pima.data[training.samples, ]
test.data <- pima.data[-training.samples, ]
```

```{r}
# Fit LDA
fit <- lda(diabetes ~., data = train.data)
# Make predictions on the test data
predictions <- predict(fit, test.data)
prediction.probabilities <- predictions$posterior[,2]
predicted.classes <- predictions$class
observed.classes <- test.data$diabetes
```

```{r}
accuracy <- mean(observed.classes == predicted.classes)
accuracy
```

```{r}
accuracy <- mean(observed.classes == predicted.classes)
accuracy

error <- mean(observed.classes != predicted.classes)
error

```

```{r}
# Confusion matrix, number of cases
table(observed.classes, predicted.classes)

# Confusion matrix, proportion of cases
table(observed.classes, predicted.classes) %>%
  prop.table() %>% round(digits = 3)
```

```{r}
confusionMatrix(predicted.classes, observed.classes,
                positive = "pos")
```
















